Title

### 2012-05-21 inc/external_forces/external_forces.tex
Constant and linear external forces

\subsubsection{Mathematical Formulation of External Force Contribution}
\ref{eq:potential_integral} is the entry point for the internal Coulombic repulsion force; since both potentials and integrals are linear, one may add additional potentials so that it becomes
\begin{equation}
  \begin{split}
    \Phi(\vec{r};t) \rightarrow & \int d\vecprime{r} \left( V(\vec{r} - \vecprime{r}) + \sum V_{ext}(\vec{r}) \right) n(\vecprime{r}; t) \\
    = & \int d\vecprime{r} V(\vec{r} - \vecprime{r}) n(\vecprime{r}; t) + \sum \int d\vecprime{r} V_{ext}(\vec{r}) n(\vecprime{r}; t) \\
    = & \Phi(\vec{r};t) + \Phi_{ext}(\vec{r};t)
  \end{split}
\end{equation}
This form of $\Phi(\vec{r};t)$ can be substituted into the definition of $K^{force}$ (\ref{eq:Kforce}), again in the basis defined in \ref{eq:coordinate_vector} 
\begin{equation}
  \begin{split}
    K^{force}_{ij} \rightarrow & \frac{1}{N} \int u_i u_j \frac{\partial f}{\partial \vec{p}} \frac{\partial}{\partial \vec{r}} \left( \Phi(\vec{r};t) + \Phi_{ext}(\vec{r};t) \right) d\vec{r} d\vec{p} \\
    \equiv & K^{int}_{ij} + \sum K^{ext}_{ij}
  \end{split}
\end{equation}
where $K^{int}$ is simply the old $K^{force}$. Noting that the spatial derivative of a potential is just the force, we can now define the matrix contributed by each external force is given by
\begin{equation}
  K^{ext}_{ij} = \frac{1}{N} \iint u_i u_j \left [ \frac{\partial}{\partial \vec{p}} f(\vec{r}, \vec{p}; t) \right ] \cdot \vec{F}_{ext}\;d^{3}x\,d^{3}p
\end{equation}
Finally, the general form of \ref{eq:gaussian_integral_dt_k} for (possibly multiple) external force ($\vec{F}_{ext}$) contributions is now
\begin{equation}
  \frac{\partial}{\partial t} A^{-1}_{ij} = K^{flow}_{ij} + K^{int}_{ij} + \sum K^{ext}_{ij}
\end{equation}

\subsubsection{Specific Forms of External Forces}
If $\vec{F}$ is a static field given by $\frac{qV}{d}\hat{z}$ then $K^{ext} = \hat{0}$. However, if the external force is a lensing field given by $\vec{F} = -M\cdot(x\hat{x}+y\hat{y})$ then 
\begin{equation}
  K^{ext} = M \cdot 
  \begin{pmatrix}
    K^{lens}_{\smallT} & 0 & 0 \\
    0 & K^{lens}_{\smallT} & 0 \\
    0 & 0 & 0
  \end{pmatrix}
  \qquad \text{where} \qquad
  K^{ext} = 
  \begin{pmatrix}
    0 & \sigma_{\smallT} \\
    \sigma_{\smallT} & 2 \gamma_{\smallT}
  \end{pmatrix}
\end{equation}
which by \ref{eq:dainvdt} yields an additional term $M_{\alpha} \sigma_{\alpha}$ which is added to the original $\frac{\partial \gamma_{\alpha}}{\partial t}$ equation.

>>>>

\subsubsection{Mathematical Formulation of External Force Contribution}
\ref{eq:potential_integral} is the entry point for the internal Coulombic repulsion force; since both potentials and integrals are linear, one may add additional potentials so that it becomes
\begin{equation}
  \begin{split}
    \Phi(\vec{r};t) \rightarrow & \int d\vecprime{r} \left( V(\vec{r} - \vecprime{r}) + \sum V_{ext}(\vec{r}) \right) n(\vecprime{r}; t) \\
    = & \int d\vecprime{r} V(\vec{r} - \vecprime{r}) n(\vecprime{r}; t) + \sum \int d\vecprime{r} V_{ext}(\vec{r}) n(\vecprime{r}; t) \\
    = & \Phi(\vec{r};t) + \Phi_{ext}(\vec{r};t)
  \end{split}
\end{equation}
This form of $\Phi(\vec{r};t)$ can be substituted into the definition of $K^{force}$ (\ref{eq:Kforce}), again in the basis defined in \ref{eq:coordinate_vector} 
\begin{equation}
  \begin{split}
    K^{force}_{ij} \rightarrow & \frac{1}{N} \int u_i u_j \frac{\partial f}{\partial \vec{p}} \frac{\partial}{\partial \vec{r}} \left( \Phi(\vec{r};t) + \Phi_{ext}(\vec{r};t) \right) d\vec{r} d\vec{p} \\
    \equiv & K^{int}_{ij} + \sum K^{ext}_{ij}
  \end{split}
\end{equation}
where $K^{int}$ is simply the old $K^{force}$. Noting that the spatial derivative of a potential is just the force, we can now define the matrix contributed by each external force is given by
\begin{equation} \label{eq:Kint}
  K^{ext}_{ij} = \frac{1}{N} \iint u_i u_j \left [ \frac{\partial}{\partial \vec{p}} f(\vec{r}, \vec{p}; t) \right ] \cdot \vec{F}_{ext}\;d^{3}x\,d^{3}p
\end{equation}
Finally, the general form of \ref{eq:gaussian_integral_dt_k} for (possibly multiple) external force ($\vec{F}_{ext}$) contributions is now
\begin{equation}
  \frac{\partial}{\partial t} A^{-1}_{ij} = K^{flow}_{ij} + K^{int}_{ij} + \sum K^{ext}_{ij}
\end{equation}

\subsubsection{Constant External Forces}

For constant static forces having no dependence in $ x $, $ y $, $ z $, or $ t $, one can readily show that there is no effect on the pulse propagation parameters $ \sigma_{ \alpha } $, $ \eta_{ \alpha } $, and $ \gamma_{ \alpha } $.
For example, consider a field consider a field $ \vec{E} = -E \hat{z} $ such that electrons are accelerated in the positive $ z $ direction by $ \vec{F}_{ext} = qE \hat{z} = q ( V / d ) \hat{z} $, so that from \ref{eq:Kint} we have
\begin{equation}
 \begin{split} 
  K^{ext}_{ij} = \frac{1} {N} \dfrac{qV} {d} \int{ u_{i} u_{j} \left( \frac{p_{z}} {\eta_{z}} - \frac{\gamma_{z} z} {\sigma_{z} \eta_{z}} \right) f d^{3}x   d^{3}p } ;
 \end{split}
\end{equation}
but it is clear that no function of this form multiplying $ f $, which is a Gaussian distribution, can avoid creating an odd integrand, and thus $ K_{ij}^{ext} = 0 $ for all $ i $ and $ j $ (or equivalently $ K^{ext} = \hat{0} $) meaning that this static field has no effect on the parameters of the pulse.
Of course, this is only true in the pulse frame, where there is no total momentum and thus $ \Delta E = \left( \Delta p \right)^{2} / 2 m_{e} $ is unchanged.
However, in the lab frame, where  $ \Delta E = p \Delta p / m_{e} $, there is an increase in the momentum spread, as the pulse now has the large momentum $p$ imparted by the accelerating field.

\subsubsection{Linear External Forces}

For the case of a perfect lens, or in the longitudinal direction a perfect RF compression element, with a parabolic potential, the external force is linear with distance from the center of the electron pulse: 
\begin{equation} \label{eq:linear_force}
\vec{F}_{ext} = - M_{ \smallT } r \hat{\operatorname{r}} - M_{z} z \hat{\operatorname{z}} \text{ .}
\end{equation}
Here, $M_{ \smallT }$ and $M_{z}$ characterize the strength of a lens or a compressor, respectively, and are chosen to be positive for pulse focusing. Substitution into the integral equation for the $ K^{ext} $ matrix elements,
\begin{equation}
 \begin{split} 
  K^{ext}_{ij} & \equiv \dfrac{1} {N} \int u_{i} u_{j} ( \vec{\nabla}_{p} f ) \cdot \vec{F}_{ext} d^{3} x d^{3} p \\ & = \dfrac{1} {N} \int u_{i} u_{j} \biggl[ x \left( \dfrac{ p_{x} } { \eta_{ \smallT } } - \dfrac{ \gamma_{ \smallT } x } { \sigma_{ \smallT } \eta_{ \smallT } } \right) + y \left( \dfrac{ p_{y} } { \eta_{ \smallT } } - \dfrac{ \gamma_{ \smallT } y } { \sigma_{ \smallT } \eta_{ \smallT } } \right) \\ & \quad + z \left( \dfrac{ p_{z} } { \eta_{z} } - \dfrac{ \gamma_{z} z } { \sigma_{z} \eta_{z} } \right) \biggr] f d^{3} x d^{3} p ,
 \end{split}
\end{equation}
leads to the $6\times6$ matrix
\begin{equation}
K^{ext} = 
\begin{pmatrix}
\hat{k}^{lin}_{\smallT} & \hat{0} & \hat{0} \\
\hat{0} & \hat{k}^{lin}_{\smallT} & \hat{0} \\
\hat{0} & \hat{0} & \hat{k}^{lin}_{z}
\end{pmatrix} \text{ ,}
\end{equation}
where
\begin{equation}
\hat{k}^{lin}_{\alpha} =
\begin{pmatrix}
0 & M_{\alpha} \sigma_{\alpha} \\
M_{\alpha} \sigma_{\alpha} & 2 M_{\alpha} \gamma_{\alpha} \text{.}
\end{pmatrix} \text{ .}
\end{equation}
which by \ref{eq:dainvdt} yields an additional term $M_{\alpha} \sigma_{\alpha}$ which is added to the original $\dfrac{d \gamma_{\alpha}}{d t}$ equation.
